\LARGE \bf TODO Name this thing

### 2012-02-18 papers/itsc_2012/root.tex
headless reports how many agents actually move during ticks, to detect deadlock. small updates to paper.

@@

% \commentd{nevada even made autonomous cars legal:
% http://www.pcmag.com/article2/0,2817,2400400,00.asp}
% This paper presents a new methodology for micro-simulating autonomous traffic for such situations where we will continue to use existing roads for autonomous vehicles. We believe that the autonomous traffic will be controlled via an overseer. The overseer can define behaviors for indivudal agents and individual intersections. The overseer itself can be a centralized or decentralized system that is capable of making such decisions in real-time.

@@

@@

\commentd{UML diagram?}

The architecture and implementation notes of AORTA follow. AORTA is divided into
roughly three components, each somewhat independent of the others. The map model
handles everything from building graphs from OSM source to pathfinding. This
component is cleanly separated from the simulation engine, which adds the notion
of agents, structures on top of map components that detect collisions, and the
machinery to implement AORTA's model of car dynamics. If a future experiment
demands macrosimulation or cellular automata method, no changes to the map
infrastructure are needed. Finally, there is a UI using Java Swing, which is
(out of necessity) tied to both the map and simulation components. The software
has a documented API, is open-source, and contains all of the code for the
experiments described in Section 6: it is designed for extensibility.

@@

Our goal is to simulate traffic on existing infrastructure, not contrived,
randomly generated maps. We do this by utilizing OpenStreetMaps, a community
project offering user-editable, free maps.

To run a simulation on a map from OSM, there is a one-time conversion process to
transform the map into AORTA's format, which is also encoded with XML. It is
split into 3 passes, each feeding the next.

@@

AORTA's map format reflects our map model, which will now be described.

\commentd{terminology sucks below. suggestions?}

\emph{Roads} are undirected edges between \emph{vertices}, which are
intersections between multiple roads. A road contains one or two sets of
directed \emph{edges}, each of which represents a single lane of travel. One-way
roads are supported. An edge leads from one vertex to another; the second vertex
then contains a mapping from each incoming edge to some outgoign edge; each pair
is a \emph{turn}. Strictly speaking, the multigraph consists of \emph{lanes} as
vertices (this is confusing, because lanes are \emph{edges} in our model) and
\emph{turns} as edges. \emph{Roads} are a super-structure with meta-data,
bundling several \emph{lanes} together, and \emph{vertices} serve to bundle
together \emph{turns}.

Geometrically speaking, a road has a sequence of points defining the line
segments constituting the center yellow line. These points could also be
interpreted as control points for a curved road, but AORTA approximates using
straight lines. Edges mimic this geometry, projecting the center line a lane's
width away. Turns are modeled with a single straight line connecting the end of
one edge to the start of another. Since both edges and turns can support
traffic, they are grouped together as \emph{traversables}, meaning they have
some sequence of line segments and a total length.

@@

The relevant parts of OSM's map model are nodes and ways. A node corresponds to
a GPS location, and ways are links between many nodes, corresponding to roads
and other paths. The first pass reads OSM's XML format, gathering ways as a
series of coordinates. Ways intersect when a node is used by multiple ways.

@@

Finally, coordinates are normalized from GPS (longitude, latitude) to a planar
coordinate system with no negative coordinates and the Y coordinate increasing
upwards. This is done for convenience of manipulation in the UI, but the scale
and offset used are retained, so a coordinate in the final form of the map can
still be mapped back to its original GPS form.

@@

OSM ways only implicitly intersect at common coordinates; AORTA's structure
demands a traditional definition of edge linking exactly two vertices. Pass 2
walks along each way's coordinates, dividing it into multiple undirected edges.

@@

@@

Next, turns are constructed at each vertex between the incoming and outgoing
edges. This is completely heuristic based: we classify a potential turn as a
left, right, or straight turn based solely on angles between edges and if the
edge's original OSM ID matches. U-turns are constructed at dead-ends. If several
lanes all cross into fewer lanes, merging is forced for the rightmost lanes.
These patterns are attempts at generalizations; this process results in many
intersections having the incorrect turns. However, without a more detailed
source, this is the best compromise. Finally, the geometry of a turn is
simplified to a single line between the ends of the edges it connects.
In the future, a more realistic curve could be drawn.

@@

AORTA has a low-granularity model of turn conflict. If two vehicles at any point
along two different turns could ever come into contact, then the turns are
always in conflict, no matter where the agent specifically is along that turn.
The alternative of tiling intersections would provide more granularity, but it
adds computational complexity and would be impractical for many of the
intersection geometries inferred from OSM.

@@

Because turns are constructed imperfectly, most large graphs have edges that are
not reachable from some portions of the graph. Simulation and pathfinding would
suffer greatly if this was the case, so pass 3 uses Tarjan's linear-time
algorithm for locating strongly connected components to remove the components of
the directed graph with the fewest members. (There is also a debugging switch to
leave these components in and highlight them in the UI.) The algorithm operates
at the granularity of lanes. If several lanes of many are disconnected within
one road, then the bad lanes are spliced out. This results in visually
unappealing geometric gaps, but if the entire road were removed, then Tarjan's
algorithm would have to be repeated to determine what edges the good lanes would
then disconnect. The end result is a connected digraph, ideal for corner
case-free pathfinding and simulation.

@@

@@

In the future, we intend to support the ability to augment an AORTA map with
data from non-OSM sources, such as information about the presence of extra lanes
for left turns. Using data such as GPS coordinates and road names, we could fuse
this extra information into an osm-based AORTA map to make simulations on that map
that much more realistic.

>>>>

\commentd{sounds fine to me, only thing is that we haven't actually focused on
  dynamic re-planning or any kind of overseer yet, although theres nothing
  stopping us from trying that idea again}

@@

\commentd{nevada even made autonomous cars legal:
http://www.pcmag.com/article2/0,2817,2400400,00.asp}

\commentd{another example long-term usage: fuel. I went to a talk by
  somebody on the engr side about powermiling and optimizing battery usage}

\commentd{we don't support lane-changing yet, and overseers arent re-implemented
yet either}

@@

\commentd{and we can do the above too. we talk about how to make the current
  behavior more flawed and humanlike in the agent reactions section.}

@@

\commentd{? UML diagram: 3 components with their capabilities listed inside, and
arrows of interactions between them?}

The architecture and implementation notes of AORTA follow. AORTA is divided into
roughly three components, each somewhat independent of the others. The map model
handles everything from building graphs from OSM source to pathfinding. This
component is cleanly separated from the simulation engine, which adds the notion
of agents, structures on top of map components that detect collisions, and the
machinery to implement AORTA's model of car dynamics. If a future experiment
demands macrosimulation or cellular automata method, no changes to the map
infrastructure are needed, due to the software engineering. Finally, there is a
UI using Java Swing, which is (out of necessity) tied to both the map and
simulation components. The software has a documented API, is open-source, and
contains all of the code for the experiments described in Section 6 -- it is
designed for extensibility.

@@

Our goal is to simulate traffic on existing infrastructure, not contrived,
randomly generated maps. We do this by utilizing OpenStreetMaps, a community
project offering user-editable, free maps.  To run a simulation on a map from
OSM, there is a one-time conversion process to transform the map into AORTA's
format, which is also encoded with XML. It is split into 3 passes, each feeding
the next.

@@

\commentd{a simple diagram of a road with a few edges connected maybe to a
one-way}

\commentd{terminology is awkward. if we bold-face our notions of edge and vertex
every time we use them, would that better distinguish them from a graph's
verts/edges?}

\emph{Roads} are undirected edges between \emph{vertices}, which are
intersections between multiple roads. A \emph{road} contains one or two sets of
directed \emph{edges}, each of which represents a single lane of travel. One-way
\emph{roads} are supported. An \emph{edge} leads from one \emph{vertex} to
another; the second \emph{vertex} then contains a mapping from each incoming
\emph{edge to some outgoing \emph{edge}; each pair is a \emph{turn}. Strictly
speaking and with confusing terminology, the multigraph consists of lanes
(\emph{edges}) as vertices and \emph{turns} as edges. \emph{Roads} are a
super-structure with meta-data, bundling several \emph{edges} together, and
\emph{vertices} serve to bundle together \emph{turns}.

Geometrically, a road has a sequence of points defining the line segments
constituting the center yellow line. These points could also be interpreted as
control points for a curved road, but AORTA approximates using straight lines.
Edges mimic this geometry, projecting the center line some multiple of a lane's
width away in the appropriate direction. Turns are modeled with a single
straight line connecting the end of one edge to the start of another (so right
turns often have $0$ length). Since both edges and turns can support traffic,
they are grouped together as \emph{traversables}, meaning they have some
sequence of line segments and a total length.

@@

The relevant parts of OSM's map model are nodes and ways. A node corresponds to
a GPS location, and ways are links between many nodes, corresponding to
driveable streets and other paths. The first pass reads OSM's XML format,
gathering ways as a series of coordinates. Ways implicitly intersect when a node
is used by multiple ways.

@@

Finally, coordinates are normalized from GPS (longitude, latitude) to a
Cartesian coordinate system with no negative coordinates and the Y coordinate
increasing upwards. This is only done for convenience of manipulation in the UI.
Any coordinate is translated back to its original GPS form when used for
computing traversable length.

@@

AORTA's structure demands a traditional definition of edge linking exactly two
vertices, rather than OSM's ways going through many common nodes. Pass 2 walks
along each way's coordinates, dividing it into multiple undirected edges.

@@

\commentd{the geometry was already explained in Map Model; I think it belongs
there if we have to cut one}

@@

Next, turns are constructed at each vertex between the incoming and outgoing
edges. This is completely heuristic based: we classify a potential turn as a
left, right, or straight turn based solely on angles between edges and if the
edge's original OSM ID matches. U-turns are constructed at dead-ends. If several
lanes all cross into fewer lanes, merging is forced for the rightmost lanes.
These patterns are attempts at generalizations; this process results in many
intersections having the incorrect turns. However, without a more detailed
source, this is the best compromise. At the time of writing, even sources like
Google Maps lack this information. (TODO how do you cite a lack of something?)
Finally, the geometry of a turn is simplified to a single line between the ends
of the edges it connects.  In the future, a more realistic curve could be drawn.

@@

AORTA has a low-granularity model of turn conflict. If two vehicles at any point
along two different turns could ever come into contact, then the turns are
always in conflict, no matter where the agent specifically is along that turn.
The alternative of tiling intersections (like \cite{JAIR08-dresner} does) would
provide more granularity, but it adds computational complexity and would be
impractical for many of the complex intersection geometries inferred from OSM.

@@

Because turns are constructed imperfectly, most large graphs have edges that are
not reachable from some portions of the graph. Simulation and pathfinding would
suffer greatly if this remained the case, so pass 3 uses Tarjan's linear-time
algorithm for locating strongly connected components to remove the components of
the directed graph with the fewest members. (There is also a debugging switch to
leave these components in and highlight them in the UI.) The algorithm operates
at the granularity of lanes. If several lanes of many are disconnected within
one road, then the bad lanes are spliced out. This results in visually
unappealing geometric gaps, but if the entire road were removed, then Tarjan's
algorithm would have to be repeated to determine what edges the good lanes would
then disconnect. The end result is a connected digraph, ideal for corner
case-free pathfinding and simulation.

@@

\commentd{this maybe doesn't belong in this paper}

@@

In the future, we intend to support the ability to augment an AORTA map with
data from non-OSM sources, such as information about the presence of extra lanes
for left turns. Using data such as GPS coordinates and road names, we could fuse
this extra information into an OSM-based AORTA map to make simulations on that map
that much more realistic.
